Cloud Simulation with the ExaHyPE-Engine\\Master Thesis Final

### 2019-01-24 writing/slides/final.tex
Small fixes for slides.

\begin{frame}{The ADER-DG Approach\footfullcite{dumbser2008unified}}
  Solve \textbf{hyperbolic conservation laws} of the form
\begin{equation}
  \label{eq:conservation-law}
 \frac{\partial}{\partial_t}  \Q + \divergence{\flux(\Q)} = \source(\bm{x}, t, \Q)
\end{equation}
with $Q$ vector of conserved variables, $\bm{x}$ position, $t$ time,  $\nabla \cdot F(Q)$ divergence of flux and $S(\bm{x}, t, Q)$ source term.

@@

\begin{frame}{The Navier-Stokes Equations (with advection coupling\footfullcite{hidalgo2011ader})}
  Vector of conserved quantities:
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right) 
\end{equation}
With $\Qrho$ density of fluid, $\Qj$ velocity density, $\QE$ energy density, $\QZ$ mass fraction of unburnt gas.

Flux:
\begin{equation}
  F(\Q, \textcolor{orange}{\gradQ}) = 
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure + \textcolor{orange}{\stressT} (\Q, \gradQ)  \\
    \Qv \cdot \left(\bm{I} \QE + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)} \right) -
    \textcolor{orange}{\kappa \nabla T}\\
    \Qj \QZZ
  \end{pmatrix}
\end{equation}

$T$ temperature, $\pressure(\Q)$ pressure,
$\textcolor{orange}{\stressT}$ stress tensor, $\textcolor{orange}{\kappa \nabla T}$ heat diffusion term
\end{frame}

@@

\textbf{Solution}:
Modify numerical flux (Riemann solver), time step size (\textsc{cfl}-Condition) and boundary conditions to \textbf{allow diffusive terms}\footfullcite{dumbser2010arbitrary}.

@@

  Total variation of $f$:
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}

@@

\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + T_\text{refine} \sigma \\
    \text{delete} & \text{if } \tv(\Q) < \mu + T_\text{delete} \sigma \\
    \text{keep} & \text{otherwise}
    \end{cases}
\end{equation}
Computation of $\mu, \sigma$ with stable, pairwise reduction\footfullcite{chan1982updating}.

@@

\begin{frame}
  \frametitle{AMR vs.\ fully refined grid: Error}
  \begin{figure}[H]
    \centering
    \ifratio{43}{
\includegraphics[width=0.9\textwidth]{thesis_two_bubbles_amr_error}
  } {
\includegraphics[width=0.6\textwidth]{thesis_two_bubbles_amr_error}
  }
\caption{Potential temperature of fully refined solution minus \amr{} grid.}
  \end{figure}
\end{frame}

@@

\begin{frame}\frametitle{Reactive Source\footfullcite{hidalgo2011ader,helzel2000modified}}
Source term for mass fraction $\QZ$:  
 \begin{equation}\label{eq:source-reaction}
  \source[\QZ] = - K(T) \QZ
\end{equation}

Reaction model:
\begin{equation}\label{eq:reaction-model}
  K(T) = 
\begin{cases}
  - \frac{1}{\reactionTimescale} & T > \reactionTemperature\\
  0 & \text{otherwise}
\end{cases}  
\end{equation}
$\reactionTimescale = 0.01$ timescale of reaction

$\reactionTemperature$ activation temperature
\end{frame}

@@

\begin{frame}{Summary}
  \begin{block}{}
  \begin{itemize}
  \item \aderdg{} \textbf{converges} for Navier Stokes equations
  \item \aderdg{} shows \textbf{good results} for large set of scenarios (many not shown here!)
  \item (Limiting) \aderdg{} works for reactive Euler \textit{\&} Navier-Stokes
  \item \amr{}-criterion is \textbf{faster} and tracks cloud \textbf{correctly}.
  \item All \textbf{extensions} to engine (diffusive fluxes, global observables) useful for \textbf{other users} of \exahype{}.
  \end{itemize}
  \end{block}

>>>>

\begin{frame}{The ADER-DG Approach\footfullcite{dumbser2008unified}}
  Solve \textbf{hyperbolic conservation laws} of the form
\begin{equation}
  \label{eq:conservation-law}
 \frac{\partial}{\partial_t}  \Q + \divergence{\flux(\Q)} = \source(\bm{x}, t, \Q)
\end{equation}
with $\Q$ vector of conserved variables, $\bm{x}$ position, $t$ time,  $\divergence{\flux(\Q)}$ divergence of flux and $\source(\bm{x}, t, \Q)$ source term.

@@

\begin{frame}{The Navier-Stokes Equations (with advection coupling\footfullcite{hidalgo2011ader})}
  Vector of conserved quantities:
\begin{equation}
  \label{eq:conserved-variables}
 \Q = \left( \Qrho, \Qj, \QE, \QZ \right) 
\end{equation}
With $\Qrho$ density of fluid, $\Qj$ velocity density, $\QE$ energy density, $\QZ$ mass fraction of chemical reactant (unburnt gas).

Flux:
\begin{equation}
  F(\Q, \textcolor{orange}{\gradQ}) = 
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)}  \\
    \Qv \cdot \left(\bm{I} \QE + \bm{I} \pressure + \textcolor{orange}{\stressT (\Q, \gradQ)} \right) -
    \textcolor{orange}{\kappa \nabla T}\\
    \Qj \QZZ
  \end{pmatrix}
\end{equation}

Pressure $\pressure(\Q)$,
stress tensor $\textcolor{orange}{\stressT}$, heat diffusion $\textcolor{orange}{\kappa \nabla T}$ with temperature $T$.
\end{frame}

@@

\textbf{Solution}:
Modify numerical flux (Riemann solver), time step size (\textsc{cfl}-Condition) and boundary conditions to \textbf{allow diffusive terms}\footfullcite{dumbser2010arbitrary,gassner2008discontinuous}.

@@

  Total variation of $\bm{x}$ for cell $\cell$:
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}

@@

\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + T_\text{refine} \sigma \\
    \text{delete} & \text{if } \tv(\Q) < \mu + T_\text{delete} \sigma \\
    \text{keep} & \text{otherwise}
    \end{cases}
\end{equation}
Computation of mean $\mu$, standard deviation $\sigma$ with \textbf{stable}, pairwise reduction\footfullcite{chan1982updating}.

@@

\begin{frame}
  \frametitle{AMR vs.\ fully refined grid: Error}
  \begin{figure}[H]
    \centering
    \ifratio{43}{
\includegraphics[width=0.9\textwidth]{thesis_two_bubbles_amr_error}
  } {
\includegraphics[width=0.6\textwidth]{thesis_two_bubbles_amr_error}
  }
\caption{Potential temperature of fully refined solution minus \amr{} solution.}
  \end{figure}
\end{frame}

@@

\begin{frame}\frametitle{Reactive Source\footfullcite{hidalgo2011ader,helzel2000modified}}
Source term for mass fraction $\QZ$:  
 \begin{equation}\label{eq:source-reaction}
  \source[\QZ] =\begin{cases}
  - \frac{1}{\reactionTimescale} \QZ  & T > \reactionTemperature\\
  0 & \text{otherwise}
\end{cases}  
\end{equation}

$\reactionTimescale = 0.01$ timescale of reaction.

$\reactionTemperature$ activation temperature.
\end{frame}

@@

\begin{frame}{Summary}
  \begin{block}{Results}
  \begin{itemize}
  \item \aderdg{} \textbf{converges} for Navier Stokes equations
  \item \aderdg{} shows \textbf{good results} for large set of scenarios (many not shown here!)
  \item \amr{}-criterion is \textbf{faster} and tracks cloud \textbf{correctly}.
  \item (Limited) \aderdg{} works for reactive Euler \textit{\&} Navier-Stokes
  \item All \textbf{extensions} to engine (diffusive fluxes, global observables) useful for \textbf{other users} of \exahype{}.
  \end{itemize}
  \end{block}
